The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-04-23 paper/ms.tex
Rephrase intro paragraph to fit new structure

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, the local midplane location
will not agree with our local midplane extrapolated onto their position,
thanks to relatively small-scale variations in the density of the gas and
stellar distribution. Converting positions and velocities of more distant
stars from a heliocentric to a Galactocentric coordinate system introduces a
systematic bias in the $z$ coordinate, and hence the actions computed for
these stars. The further from the solar neighborhood the target star, the more
likely the mismatch will result in large systematic uncertainty, especially in
the vertical action $J_z$. Depending on the type of orbit being studied, this
sets a limit on the ``dynamical solar neighborhood'' over which the local
coordinate frame can be usefully extended for action-space analysis. A similar
argument applies to any remaining uncertainty in measurements of the
Galactocentric radius of the Sun $R_{\odot}$ and to variations in the local
standard of rest (LSR).

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions. Lacking an empirical, nonlocal
definition of the global Galactic midplane or a measurement of the expected
midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
we examine the expected azimuthal dependence of the local midplane, and hence
the expected size of the dynamical solar neighborhood, using cosmological,
hydrodynamical, zoom-in simulations of Milky Way-mass galaxies from the
Feedback in Realistic Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}}
\citep{2016ApJ...827L..23W, 2017arXiv170206148H}.

>>>>

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, the local midplane location
will not agree with our local midplane extrapolated onto their position,
due to relatively small-scale variations in the density of the gas and
stellar distribution. Converting positions and velocities of more distant
stars from a heliocentric to a Galactocentric coordinate system introduces a
systematic bias in the $z$ coordinate, and hence the actions computed for
these stars. The further from the solar neighborhood the target star, the more
likely the mismatch will result in large systematic uncertainty, especially in
the vertical action $J_z$. Depending on the type of orbit being studied, this
sets a limit on the ``dynamical solar neighborhood'' over which the local
coordinate frame can be usefully extended for action-space analysis. A similar
argument applies to any remaining uncertainty in measurements of the
Galactocentric radius of the Sun and to variations in the local
standard of rest (LSR).

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions. We then discuss two potential
sources of these coordinate system offsets in
Sections~\ref{sec:mes_err}~and~\ref{sec:local_file}. In
Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
parameters of the Galactocentric coordinate system. In
Section~\ref{sec:local_fire} we examine the azimuthal variations of the
midplane itself in two sets of simulations: cosmological, hydrodynamical,
zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}}
\citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
simulation of a Sagittarius encounter with a Milky Way-mass galaxy
\citep{2018MNRAS.481..286L}.
